### 2014-08-10 inst/simulation/manuscript.Rmd
...

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
scenarios <- c("A","B","C","D","E","F")
modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"))

@@

```{r track-data}
#scenario <- "A"
#iterations <- as.integer(1)

iterations <- as.integer(c(1,1,1,1,3,1))
i <- 1
trackData <- data.frame()
surveyRoutesData <- data.frame()

@@

p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z)) + scale_fill_manual(values=col) +
  geom_path(data=trackData, aes(long, lat, group=group, colour=observation)) + scale_colour_manual(values=c("darkgray","black")) +
  geom_path(data=surveyRoutesData, aes(long, lat, group=group)) +
  coord_fixed(ratio=1, xlim=c(3.375,3.4)*1e6, ylim=c(6.9,6.925)*1e6) +
  facet_wrap(~scenario) + #coord_equal()
  theme_raster() + theme(panel.border=element_rect(fill="transparent"))
print(p)
saveFigure(p, "manuscript-figure-1.png", width=8, height=4)
```

```{r figure-2}
# TODO: add densities

surveyLocations <- ddply(surveyRoutesData, .(group, scenario), function(x) data.frame(long=mean(x$long), lat=mean(x$lat), scenario=x$scenario[1], observed=x$observed[1]))[,c("long","lat","scenario","observed")]
surveyLocations$variable <- "Survey routes"

trueLocations <- subset(trackData, observation==T)
trueLocations <- ddply(trueLocations, .(burst, scenario), function(x) data.frame(long=mean(x$long), lat=mean(x$lat), scenario=x$scenario[1], observed=F))[,c("long","lat","scenario","observed")]
trueLocations$variable <- "Individual locations"

x <- rbind(surveyLocations, trueLocations)

template <- raster(extent(study$studyArea$boundary), nrows=12, ncols=6, crs=study$studyArea$proj4string)
template@extent@xmax <- template@extent@xmin + dim(template)[2] * 100 * 1e3
template@extent@ymax <- template@extent@ymin + dim(template)[1] * 100 * 1e3

p <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey") +
  geom_point(data=x, aes(long, lat, colour=observed)) + scale_color_manual(values=c("darkgrey", "black")) + 
  facet_grid(variable~scenario) + theme_raster() + coord_equal() +
  scale_x_continuous(breaks=seq(template@extent@xmin, template@extent@xmax, 1e5)) +
  scale_y_continuous(breaks=seq(template@extent@ymin, template@extent@ymax, 1e5)) +
  theme(panel.grid.major=element_line(colour="grey"))
print(p)
saveFigure(p, "manuscript-figure-2.png", width=8, height=4)
```

```{r figure-3}
validation <- Validation()
populationSizes.temporal <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)

p <- ggplot(populationSizes.temporal, aes(Observed, Estimated)) + geom_point() + stat_smooth(method="lm", colour="black") +
  geom_abline(colour="darkgrey") + facet_wrap(~Scenario, scales="free") +
  theme_minimal() + theme(panel.border=element_rect(fill="transparent"))
print(p)
saveFigure(p, "manuscript-figure-3.png", width=8, height=4)
```

```{r figure-4}
validation <- Validation()
populationSizes.spatial <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames)

@@

>>>>

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
scenarios <- c("A","B","C","D","E","F")
modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"))
iterations <- as.integer(c(1,1,1,1,3,1))

@@

```{r figure-1-init}
#scenario <- "A"
#iterations <- as.integer(1)

i <- 1
trackData <- data.frame()
surveyRoutesData <- data.frame()

@@

p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z)) + scale_fill_manual(values=col) +
  geom_path(data=trackData, aes(long, lat, group=group, colour=observation)) +#, alpha=observation)) +
  scale_colour_manual(values=c("gray40","black")) +# scale_alpha_manual(values=c(.5,1)) +
  geom_path(data=surveyRoutesData, aes(long, lat, group=group)) +
  coord_fixed(ratio=1, xlim=c(3.375,3.4)*1e6, ylim=c(6.9,6.925)*1e6) +
  facet_wrap(~scenario) + #coord_equal()
  theme_raster(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
#print(p)
saveFigure(p, "manuscript-figure-1.png", width=8, height=4)

getRaster <- function(habitatRaster, xlim, ylim, size=10000) {
  x <- crop(habitatRaster, extent(c(xlim, ylim)))
  y <- sampleRegular(x, size=size, asRaster=TRUE)
  habitatData <- as.data.frame(rasterToPoints(y))
  colnames(habitatData) <- c("x","y","z")
  habitatData$z <- factor(habitatData$z, levels=0:255)
  col <- habitatRaster@legend@colortable
  names(col) <- 0:255
  return(habitatData)
}

getPanel <- function(scenario, xlim=c(3.375,3.4)*1e6, ylim=c(6.9,6.925)*1e6, habitatRaster, trackData, surveyRoutesData) {
  s <- scenario
  p <- if (!missing(habitatRaster)) {
    habitatData <- getRaster(habitatRaster=habitatRaster, xlim=xlim, ylim=ylim)
    ggplot(habitatData) + geom_raster(aes(x, y, fill=z)) + scale_fill_manual(values=col) +
      geom_path(data=subset(trackData, scenario==s), aes(long, lat, group=group, colour=observation))
  }
  else {
     ggplot(subset(trackData, scenario==s)) +
       geom_path(aes(long, lat, group=group, colour=observation))
  }

  p <- p + scale_colour_manual(values=c("gray40","black")) +
    geom_path(data=subset(surveyRoutesData, scenario==s), aes(long, lat, group=group)) +
    coord_fixed(ratio=1, xlim=xlim, ylim=ylim) +
    theme_raster(base_size=10, panel.border=element_rect(fill="transparent"), plot.title=element_text(size=12)) +
    ggtitle(s)

  return(p)
}

library(gridExtra)

p1 <- getPanel(scenario="A", trackData=trackData, surveyRoutesData=surveyRoutesData)
p2 <- getPanel(scenario="B", trackData=trackData, surveyRoutesData=surveyRoutesData)
p3 <- getPanel(scenario="C", trackData=trackData, surveyRoutesData=surveyRoutesData)
p4 <- getPanel(scenario="D", trackData=trackData, surveyRoutesData=surveyRoutesData)
p5 <- getPanel(scenario="E", habitatRaster=study$studyArea$habitat, trackData=trackData, surveyRoutesData=surveyRoutesData)
p6 <- getPanel(scenario="F", habitatRaster=study$studyArea$habitat, trackData=trackData, surveyRoutesData=surveyRoutesData)
p <- arrangeGrob(p1, p2, p3, p4, p5, p6, ncol=3, nrow=2)
saveFigure(p, "manuscript-figure-1.png")
```

```{r figure-2-init}
getSurveyRouteLocations = function(scenarios, modelName, iterations, years) {
  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, years) {
    message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
    study <- getStudy(scenario=x$scenario)
    estimates <- SimulatedSmoothModelSpatioTemporal(study=study, iteration=x$iteration)
    estimates$modelName <- modelName
    estimates <- study$loadEstimates(estimates)
    surveyRoutes <- data.frame(estimates$getUnscaledObservationCoordinates(), Intersection=subset(estimates$data, year %in% years)$intersections>0)
    surveyRoutes$Scenario <- x$scenario
    surveyRoutes$Iteration <- x$iteration
    return(surveyRoutes)
  }, modelName=modelName, years=years)
}

getAgentLocations = function(scenarios, modelName, iterations, years) {
  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, years) {
    study <- getStudy(scenario=x$scenario)
    tracks <- study$loadTracks(iteration=x$iteration)
    tracks$tracks <- subset(tracks$tracks, year %in% years & yday<59)
    intersections <- study$loadIntersections(iteration=x$iteration)

    # If no intersection, randomized track can be anywhere else except on the survey route

    observedBursts <- aaply(intersections$intersectionsMatrix, 2, function(x) if (any(x>0)) TRUE else FALSE)
    observationTracks <- tracks$randomizeObservationDayTracks(days=1)

    a <- ddply(observationTracks$tracks, .(id, year, burst), function(x, observationTracks) {
      name <- with(x[1,], paste(id, year, burst))
      if (observedBursts[name] == FALSE) return(data.frame(x=mean(x$x), y=mean(x$y), year=x$year))
      NULL
    }, observationTracks=observationTracks)[,c("x","y","year")]
    a$Observed <- F

    index <- intersections$intersections$intersections > 0 & intersections$intersections$year %in% years
    if (all(index == FALSE)) b <- data.frame() else {
      x <- intersections$intersections[index,]
      b <- data.frame(coordinates(x), year=x$year)
      b$Observed <- T
    }
    agentLocations <- rbind(a, b)
    agentLocations$Scenario <- x$scenario
    agentLocations$Model <- x$modelName

    return(agentLocations)
  }, modelName=modelName, years=years)
}

getEstimatedDensity = function(scenarios, modelName, iterations, year, sd=FALSE) {
  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, year0, sd) {
    message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
    study <- getStudy(scenario=x$scenario)
    estimates <- SimulatedSmoothModelSpatioTemporal(study=study, iteration=x$iteration)
    estimates$modelName <- modelName
    estimates <- study$loadEstimates(estimates)
    estimates$collectEstimates()

    estimates$data <- subset(estimates$data, year == year0)
    density <- estimates$getPopulationDensity(getSD=sd)
    density <- if (sd) density$sd else density$mean
    if (is.null(density) || maxValue(density$rasterStack[[1]]) > 1) stop("Estimation failed.")
    density <- as.data.frame(density$rasterStack[[1]], xy=T)
    density$Scenario <- x$scenario
    density$Iteration <- x$iteration
    density$Model <- modelName
    density$Year <- year0

    return(density)
  }, modelName=modelName, year0=year, sd=sd)
}
```

```{r figure-2}
iterations <- as.integer(c(1,1,3,2,5,6))
#iterations <- as.integer(c(1,1,3,2,5,7))
surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios, modelName=modelNames[1], iterations, years=2001)
surveyRoutes <- rbind(surveyRoutes[surveyRoutes$Intersection == FALSE,], surveyRoutes[surveyRoutes$Intersection == TRUE,])
agents <- getAgentLocations(scenarios=scenarios, modelName=modelNames[1], iterations, years=2001)
densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=2001, sd=F)
densitiesScaled <- ddply(densities, .(scenario), function(x) { x$X2001 <- scale(x$X2001); return(x) })
limits <- extent(study$studyArea$boundary)

#p <- ggplot(boundaryDF) + geom_polygon(aes(long, lat, group=group), fill=terrain.colors(100)[1]) +
p <- ggplot(densitiesScaled) + geom_raster(aes(x, y, fill=X2001)) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black")) + 
  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal()
print(p)
saveFigure(p, "manuscript-figure-2a.png", width=8)

gridTemplate <- raster(extent(study$studyArea$boundary), nrows=12, ncols=6, crs=study$studyArea$proj4string)
gridTemplate@extent@xmax <- gridTemplate@extent@xmin + dim(gridTemplate)[2] * 100 * 1e3
gridTemplate@extent@ymax <- gridTemplate@extent@ymin + dim(gridTemplate)[1] * 100 * 1e3

p <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey") +
  geom_point(data=surveyRoutes, aes(x, y, colour=Intersection), size=1) + scale_color_manual(values=c("darkgrey", "black")) +
  scale_x_continuous(breaks=seq(gridTemplate@extent@xmin, gridTemplate@extent@xmax, 1e5), expand=c(0,0)) +
  scale_y_continuous(breaks=seq(gridTemplate@extent@ymin, gridTemplate@extent@ymax, 1e5), expand=c(0,0)) +
  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal() + theme(panel.grid.major=element_line(colour="grey"))
print(p)
saveFigure(p, "manuscript-figure-2b.png", width=8)
```

```{r figure-3-test, eval=FALSE}
study <- getStudy(scenario="A")
validation <- Validation(study=study)
x <- validation$validateTemporalPopulationSize(modelNames[1])
validation$summarizePopulationSize(x)
```

```{r figure-3}
study <- getStudy(scenario="A")
validation <- Validation(study=study)
populationSizes <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)
p <- ggplot(populationSizes, aes(Observed, Estimated, colour=modelName)) + geom_point() + stat_smooth(method="lm", colour="black") +
  scale_colour_manual("Model", values=c("black", "darkgray"), labels=c("ST", "T")) +
  geom_abline(colour="darkgrey", linetype="longdash") + facet_wrap(~Scenario, scales="free") +
  xlab("True population size") + ylab("Estimated population size") +
  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
print(p)
saveFigure(p, "manuscript-figure-3.png", width=8, height=4)
```

```{r figure-4-test, eval=FALSE}
study <- getStudy(scenario="F")
validation <- Validation(study=study)
x <- validation$validateSpatialPopulationSize(modelNames[1])
```

```{r figure-4}
study <- getStudy(scenario="A")
validation <- Validation(study=study)
populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames)
p <- ggplot(populationCorrelations, aes(True, Correlation, colour=modelName)) + geom_point() +
  scale_colour_manual("Model", values=c("black", "darkgray"), labels=c("ST", "T")) +
  facet_wrap(~Scenario) + xlab("True population size") +
  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
print(p)
saveFigure(p, "manuscript-figure-4.png", width=8, height=4)
```

```{r figure-5-init}

@@

```{r figure-5}
study <- getStudy(scenario="A")
validation <- Validation(study=study)
a <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames)
b <- validation$getValidatedCredibilityIntervalsProportions(modelNames=modelNames, probs=c(.25, .75), probsName="50%")
credibilityIntervals <- rbind(a, b)
p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Interval, colour=modelName)) + geom_point() +
  scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("black", "darkgray"), labels=c("ST", "T")) +
  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
  xlab("Scenario")
print(p)
saveFigure(p, "manuscript-figure-5.png", width=4, height=2)
```

### 2014-08-10 inst/wtc-finland/manuscript.Rmd
...

>>>>

```{r gps-stats}
library(plyr)
response <- "canis.lupus"

study <- FinlandWTCStudy$new(context=context, response=response)
tracks <- FinlandWTCTracks$new(study=study)
tracks$loadTracks()
tracks$loadMetadata()
xtabs(sex~born, tracks$metadata)

library(reshape2)

dcast(tracks$metadata, .~sex)
dcast(tracks$metadata, .~born)
dcast(tracks$metadata, .~dispersing)$"TRUE"
dcast(tracks$metadata, .~homerange)$"TRUE"
dcast(tracks$metadata, .~alpha)

x <- ld(tracks$tracks)$dt/3600
sort(x[x<=24])
sort(x[x<=24])*3600

tracksSP <- tracks$getSpatialLines(variables=.(id,year,burst))
x <- tracksSP@data
x$juvenile <- if ("born" %in% colnames(x)) factor(ifelse(x$year - x$born <= 1, "juvenile", "adult")) else factor("unknown")
x$juvenile <- factor(x$juvenile, levels=c(levels(x$juvenile), "unknown"))
x$juvenile[is.na(x$juvenile)] <- "unknown"
dcast(x, .~juvenile)
